### 2016-02-08 Manuscript/SW_early_life_telomeres.Rmd
More edits

#####Early-life telomere length and age
Inter-plate repeatability of telomere length, based on XX samples measured at least twice, was 0.78 (CI = 0.73-0.82). Telomere length dereased with age within the first year of life (R^2^ = `r round2(summary(lm(LogTL~Agemonths,data=juv))$r.squared)`; F = `r round2(anova(lm(LogTL~Agemonths,data=juv))[1,4])`; P = `r round3(anova(lm(LogTL~Agemonths,data=juv))[1,5])`; Fig. 1A). Similarly, telomere shortening based on longitudinal data was significantly greater than zero in nestlings (one sample t-test; P < 0.001), but not in any other age group (all P > 0.5; Fig. 1B).

@@

#####Individual-level variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class and the number of non-helping subordinates in the natal territory (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lmer(LogTL~1+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE))-AICc(lmer(LogTL~Fledged+NonHelper+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE)))`). The age class effect was very strong and significant, and was present in every model in the top model set (Fig. 3A). The effect of non-helpers, on the other had did not have higher importance across the top model set than the other variables, and confidence intervals overlapped zero (Fig. 3A). Furthermore the model including non-helpers and age class model was not a significantly better fit than a model including age class alone ($\Delta$AICc = 0.06; Table S1).

>>>>

#####Early-life telomere dynamics
Inter-plate repeatability of telomere length, based on XX samples measured at least twice, was 0.78 (CI = 0.73-0.82). Telomere length dereased with age within the first year of life (R^2^ = `r round2(summary(lm(LogTL~Agemonths,data=juv))$r.squared)`; F = `r round2(anova(lm(LogTL~Agemonths,data=juv))[1,4])`; P = `r round3(anova(lm(LogTL~Agemonths,data=juv))[1,5])`; Fig. 1A). Longitudinal data showed that within-individual telomere length between early and later life samples was significantly, albeit weakly, correlated (R^2^ = `r round2(summary(lm(LogTL1~LogTL,data=Loss))$r.squared)`; F = `r round2(anova(lm(LogTL1~LogTL,data=Loss))[1,4])`; P = `r round3(anova(lm(LogTL1~LogTL,data=Loss))[1,5])`; Fig. 1B). In nestlings, the absolute reduction telomere in telomere length was not related to the length of time between sampling events (R^2^ = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 1)))$r.squared)`; F = `r round2(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 1)))[1,4])`; P = `r round3(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 1)))[1,5])`), but was consistently greater than zero, even when the second sample was taken very shortly after the first (Fig. 1C). Conversely, in both fledglings and subadults, the reduciton in telomere length was, on average, close to zero when the second sample was taken shortly after the first. The reduction in telomere length increased with the amount of time between sampling events in subadults (R^2^ = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 10)))$r.squared)`; F = `r round2(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 10)))[1,4])`; P = `r round3(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 10)))[1,5])`), but not fledglings (R^2^ = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 6)))$r.squared)`; F = `r round2(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 6)))[1,4])`; P = `r round3(anova(lm(Loss~TimeDiff,data=subset(Loss,Agemonths == 6)))[1,5])`; Fig. 1C). Conversely, the rate of telomere shortening decreased with the time between sampling events in chicks (R^2^ = `r round2(summary(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 1)))$r.squared)`; F = `r round2(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 1)))[1,4])`; P = `r round3(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 1)))[1,5])`), but not in fledglings (R^2^ = `r round2(summary(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 6)))$r.squared)`; F = `r round2(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 6)))[1,4])`; P = `r round3(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 6)))[1,5])`) or subadults (R^2^ = `r round2(summary(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 10)))$r.squared)`; F = `r round2(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 10)))[1,4])`; P = `r round3(anova(lm(TROC~TimeDiff,data=subset(Loss,Agemonths == 10)))[1,5])`). Together, these results suggest that in our samples, telomere shortening occurs at the greatest rate shortly after the nestling stage, and comparitively little telomere shortening occurs later in life.

@@

#####Early-life environment and early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age, territor the number of non-helping subordinates in the natal territory (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lmer(LogTL~1+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE))-AICc(lmer(LogTL~Fledged+NonHelper+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE)))`).

### 2016-02-08 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More edits

Fig1D <- ggplot(temp,aes(x = TimeDiff,y = TROC,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss rate (kb/year)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 8,label = 'D')

>>>>

Fig1D <- ggplot(temp,aes(x = TimeDiff,y = TROC,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss rate (kb/year)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  stat_smooth(method = 'lm', formula = y~(1/exp(x)))+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 8,label = 'D')
